# Quality of Red Wine 

## Haakon Troemborg

### 15th of October 2015

========================================================

```{r global_options, include=FALSE}
  knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(plyr)
library(ggplot2)
library(GGally)
library(scales)
library(memisc)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
setwd("C:/Users/hakon.tromborg/Data Analyst Nanodegree/R")
rw <- read.csv('./project/wineQualityReds.csv')
```

```{r}
str(rw)
```
For each wine, there are 10 different observations and one index (X).
```{r}
summary(rw)
```

Even though the quality is on a scale from 0 to 10, the wines are mostly either 5 or 6, and minimum 3 and maximum 8. This might have something to do with the fact that wine tasting is somewhat subjective, and the score is the median of several wine experts. All the data in the set is numeric.




# Univariate Plots Section
```{r echo=FALSE}
#Quality
ggplot(aes(x=quality),data=rw) + 
  geom_histogram(binwidth=1)+
  scale_x_discrete(breaks=seq(2,9,1))
```

The quality seems to be approximately normally distributed around 5.5. Interestingly, it doesn't utilize the full scale of 0 to 10. It's gonna be interesting to see how this plays out compared to diamonds, where the price was more of a continuous nature.

```{r echo=FALSE}
#Alcohol
ggplot(aes(x=alcohol),data=rw) + 
  geom_histogram(binwidth=0.1)+
  scale_x_continuous(limits=c(8,15),
                     breaks=seq(8,15,0.5))
```

Alcohol seems to have more of a skewed distribution, where most of the wines are somewhere in the range of 9 to 12%. Alcohol percentage is something that many people claim to be able to taste, it's gonna be interesting to see how this influences the quality.

```{r echo=FALSE}
#Residual sugar
ggplot(aes(x=residual.sugar),data=rw) + 
  geom_histogram(binwidth=0.1)+
  scale_x_continuous(limits=c(1,5),
                     breaks=seq(1,5,0.5))

summary(rw['residual.sugar'])
```

Almost all of the sugar levels are in the range 1 to 3, but the other values can sometimes be much higher. I'll have a look at these outliers in particular later on, to see if the they show some pattern in quality.

```{r echo=FALSE}
#Fixed acidity
ggplot(aes(x=fixed.acidity),data=rw) + 
  geom_histogram(binwidth=0.3)+
  scale_x_continuous(limits=c(4,16),
                     breaks=seq(4,16,1))
```

```{r echo=FALSE}
#Volatile acidity
ggplot(aes(x=volatile.acidity),data=rw) + 
  geom_histogram(binwidth=0.03)+
  scale_x_continuous(limits=c(0,1.6),
                     breaks=seq(0,1.6,0.25))
```

```{r echo=FALSE}
#Citric acid
ggplot(aes(x=citric.acid),data=rw) + 
  geom_histogram(binwidth=0.01)+
  scale_x_continuous(limits=c(0,1),
                     breaks=seq(0,1,0.25))
```

It's hard to believe that the wine tasters are able to taste the difference between three acids, and so I'll look into correlations between the three, and also how they relate to quality on their own and in common. Acids can obviously correlate with other qualities in wine though, so I'm not sure what to expect.
Citric acid has two spikes, at 0 and 0.5. I speculate that thisis due to rounding error at 0.5, and the fact that some wines might contain almost no citric acid.

```{r echo=FALSE}
#Chlorides
ggplot(aes(x=chlorides),data=rw) + 
  geom_histogram(binwidth=0.003)+
  scale_x_continuous(limits=c(0,0.2),
                     breaks=seq(0,0.2,0.02))
```

Chloride distribution looks pretty neat and normal, and saltiness is something I believe humans can taste pretty well. Beginning to notice a trend in histograms where the outliers are mostly on the higher end.

```{r echo=FALSE}
#Free sulfur dioxide
ggplot(aes(x=free.sulfur.dioxide),data=rw) + 
  geom_histogram(binwidth=1)+
  scale_x_continuous(limits=c(0,100),
                     breaks=seq(0,100,5))

#Total sulfur dioxide
ggplot(aes(x=total.sulfur.dioxide),data=rw) + 
  geom_histogram(binwidth=2)+
  scale_x_continuous(limits=c(0,200),
                     breaks=seq(0,200,10))
```

The distribution of the ammount of sulfur dioxide in the wine is very heavy tailed on the left side, so let's have a look at its log transform.

```{r echo=FALSE}
ggplot(aes(x=log(total.sulfur.dioxide)),data=rw) + 
  geom_histogram(binwidth=0.1)+
  scale_x_continuous()
```

Seems to be more normal now. Once again, I believe these two will be related just like the acids. Total sulfur dioxide of course contains all the free dioxide, and we'll see how they differ later on.

```{r echo=FALSE}
#pH
ggplot(aes(x=pH),data=rw) + 
  geom_histogram(binwidth=0.02)+
  scale_x_continuous(limits=c(3,5),
                     breaks=seq(3,5,0.1))
```

Sourness, like sweetness and saltiness, should be something that influences the taste of the win. I expect this to be a significant factor towards quality.

```{r echo=FALSE}
#Sulphates
ggplot(aes(x=sulphates),data=rw) + 
  geom_histogram(binwidth=0.04)+
  scale_x_continuous(limits=c(0,2),
                     breaks=seq(0,2,0.4))
```

I remember from chemistry that sulfates typically taste and smell very strongly. The amount seems fairly normally distributed, so it'll be interesting to have a look at the correlation as well as the outliers here.

# Univariate Analysis

### What is the structure of your dataset?
There are 1599 observations of 13 variables. Almost all are different measurements of chemical components in red wine, given as different weights per volume, or percentage of total volume. These are in principal continuous, but behave discretely with a relateively high resolution because of rounding and measurement accuracy. Quality stands out, which is given in integers from 3 to 8.

### What is/are the main feature(s) of interest in your dataset?
The main feature is quality, which is a mean of several expert wine tasters' opinions of the different wines. The other features are equal, what will decide interest here is whether they show particular correlations with quality.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
As stated previously, I'm not looking to decide which features are "most interesting" yet. If I were to make a guess, I'd say chlorine, sugar, alcohol, sulphate and citric acid are the most distinguishable tastes though.

### Did you create any new variables from existing variables in the dataset?
No.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?
The distributions are in general either approximately normal, or long-tailed distributions, with outliers on the upper side.
I didn't perform any operations on the data.


# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}
round(cor(rw),2)
```

We see that some of the values correlate pretty strongly, like citric and Volatile acidity, or fixed acidity and density. Let's have a look at the ones that correlate most with quality.

```{r echo=FALSE}

ggpairs(rw,columns=c(13,3,9,10,12), 
        title = "Pair plot", 
        params = c(shape = I('.'), outlier.shape = I('.')),corSize=8) + 
  ggplot2::theme_bw()
```



This plot gives a lot of information, and I'll explore some of it further instead of discussing it all at once, let's have a closer look at some of the scatter plots.

```{r echo=FALSE}
ggplot(aes(x=quality,y=residual.sugar),data=rw) + 
  geom_point(position="jitter",
             alpha = 0.7,
             color=I("Red"))
```

Nearly all wines with the higher sugar ammounts are quality 5, 6 or 7. It's easy to get tricked here, but notice that most wines in total are also 5,6 or 7. There's a qualty 4 wine with a really high ammount of sugar as welll, and it's not easy to conclude whether what we're seeing is some sort of correlation, or simply the fact that there are so many more wines with quality 5 or 6.

```{r echo=FALSE}
ggplot(aes(factor(quality),residual.sugar),data=rw) + 
  geom_boxplot() +
  scale_y_continuous(limits=c(1,3))
```

Looking at the box plot, we do see that qualities 3,4 and 8 have slightly lower means than 5,6 and 7, but nothing major.

```{r echo=FALSE}
ggplot(aes(x=quality,y=alcohol),data=rw) + 
  geom_point(position="jitter",
             alpha = 0.7,color=I("Blue"))

```

In this plot, we see more of a trend. It looks as if, in general, high quality wines have a bit more alocohol.

```{r echo=FALSE}
ggplot(aes(factor(quality),alcohol),data=rw) + 
  geom_boxplot() +
  scale_y_continuous(limits=c(8,14))
```

The box plot makes it even more obvious, especially from 5 through 8 there's a higher ammount of alcohol in the better wines. Let's look at the correlations both with and without the lower quality wines.
```{r}
cor(rw[,c("alcohol","quality")])

cor(subset(rw, quality>4)[,c("alcohol","quality")])
```

As we can see, the correlation factor is quite high, and actually quite a lot higher when we remove the quality 3 and 4 wines.

```{r echo=FALSE}
ggplot(aes(x=quality,y=chlorides),data=rw) + 
  geom_point(position="jitter",
             alpha = 0.7,color=I("Purple"))

```

It's the same thing with this plot as with the sugar one, the differences we observe are possibly just due to a higher number of quality 5 and 6 wines. 


```{r echo=FALSE}
ggplot(aes(factor(quality),chlorides),data=rw) + 
  geom_boxplot() +
  scale_y_continuous(limits=c(0,0.2))
```


```{r echo=FALSE}
ggplot(aes(x=quality,y=pH),data=rw) + 
  geom_point(position="jitter",
             alpha = 0.7,color=I("Orange"))

```

This has a different distribution than the others, and if we scroll back up to look at the histogram for pH, we see that it's normally distributed in contrary to the first three variables we've just looked at which are all heavy sided on the left. 

```{r echo=FALSE}
ggplot(aes(factor(quality),pH),data=rw) + 
  geom_boxplot() +
  scale_y_continuous(limits=c(2.8,3.75))
```

Looking at the boxplot, there seems to be a trend towards higher quality meaning also lower pH. However, the correlation factor between the two is only -0.0577.


```{r echo=FALSE}
ggplot(aes(x=quality,y=sulphates),data=rw) + 
  geom_point(position="jitter",
             alpha = 0.7,color=I("Black"))
```

We see some sort of linear trend again, for example not a single one of the really good wines have a low ammount of sulphates.

```{r echo=FALSE}
ggplot(aes(factor(quality),sulphates),data=rw) + 
  geom_boxplot() +
  scale_y_continuous(limits=c(0.25,1.3))
```

The boxplot tells the same story!

```{r echo=FALSE}
ggplot(aes(x=quality,y=volatile.acidity),data=rw) + 
  geom_point(position="jitter",
             alpha = 0.7,color=I("Red"))
```

```{r echo=FALSE}
ggplot(aes(factor(quality),volatile.acidity),data=rw) + 
  geom_boxplot() +
  scale_y_continuous(limits=c(0.2,1.2))
```

```{r echo=FALSE}
ggplot(aes(x=fixed.acidity,y=volatile.acidity),data=rw) + 
  geom_point(position="jitter")
```

Here we see that the wines with the highest ammount of fixed acidity in general don't have as high ammount of volatile acidity as the rest of the wines. However, we also have a bigger span in the quantiles here.

```{r echo=FALSE}
ggplot(aes(x=factor(round_any(fixed.acidity,2.5)),volatile.acidity),data=rw) + 
  geom_boxplot() + 
  scale_y_continuous(limits=c(0,1)) +
  xlab("fixed.acidity (grouped)")
```


```{r echo=FALSE}
ggplot(aes(x=citric.acid,
           y=fixed.acidity,
           color = factor(quality)),
           data=rw) + 
  geom_point() + 
  scale_x_continuous(limits=c(0,0.75),
                     breaks=seq(0,0.75,0.1)) +
  scale_colour_brewer(palette = "YlOrRd")
```

This trend is clearer: there's some relationship between fixed and citric acid. It could be approximately linear, or one could use a polynomial  or exponential function.

```{r echo=FALSE,message = FALSE}

ggplot(aes(x=citric.acid,
           y=fixed.acidity,
           color = factor(quality)),
           data=rw) + 
  geom_point() + 
  scale_colour_brewer(palette = "YlOrRd") + 
  stat_smooth(method="lm",
              se = FALSE,
              size=1,color = "green") + 
  stat_smooth(method="lm", 
              formula= y ~ poly(x,2) ,
              se = FALSE,
              color = "black",size = 1) + 
  scale_x_continuous(limits=c(0,0.75),breaks=seq(0,0.75,0.05))
```


Here's a look at a linear (green) and a quadratic (black) trendline.

```{r echo=FALSE}
ggplot(aes(x=free.sulfur.dioxide,y=total.sulfur.dioxide),data=rw) + 
  geom_point(position = "jitter",
             alpha=0.5) + 
  geom_smooth(method="lm")
```

Again there's a clear trend, but this is rather obvious since the two axes are total and free sulfur dioxide. Total sulfur dioxide consists of the free plus the bound sulfur, and so naturally there's a relationship between the two, most likely displaying some sort of linearity plus noise.

```{r echo=FALSE}
ggplot(aes(x=residual.sugar,y=density),data=rw) + 
  geom_point()
```

Something seems to be going on here, but the outliers of sugar are messing things up. Let's remove them!

```{r echo=FALSE}
ggplot(aes(x=residual.sugar,y=density),
       data=subset(rw, residual.sugar<quantile(residual.sugar,0.90))) + 
  geom_point(position = "jitter")
```

Here I've taken away 10% of the wines, the ones with the lowest ammount of sugar. There's clearly a trend between the two, which I guess can be attributed to the fact that sugar levels affect density in fluids.

Let's look at the correlations. Here's what it used to be:
```{r}
cor(rw[,c("residual.sugar","density","quality")])
```

And here's the 90% with the lowest sugar levels:
```{r}
cor(subset(rw, residual.sugar<quantile(residual.sugar,0.90))
    [,c("residual.sugar","density","quality")])
```



We see the correlation between sugar and density increase from 0.355 to 0.423, which is quite significant. Also notice correlations going (slightly) up for quality and both density and residual sugar. This could indicate that the outliers in terms of sugar aren't following the same rules as other wines.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?
The data sets contain several correlating variables, but the correlations are usually not that big (>0.7). Quality clearly varied with especially alcohol, volatile acidity and sulphates. 

Quality and alcohol correlate somewhat linearly, and it's worth noting that the difference is especially big on the higher qualities.

We see the same trend for sulphates, and as with alcohol the trend is clearer in the higher range. The samples with quality 3 seem to be somewhat of an outlier: many times they aren't really following the trend. This could be because there are just a few samples of quality 3, or because of some other trait that is special for quality 3 wines.


### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?
I expected some features to correlate because of their chemical nature. For example the total ammount of sulfuric oxide contains all the free oxide, and so naturally they would correlate. 

Fixed acidity seemed to decrease with volatile acidity as well.

Sugar was shown to correlate quite strongly with density for the lower 90% of sugar levels. Removing the top 10% sugar also increased correlations between quality and both sugar and density. 

### What was the strongest relationship you found?

The strongest relationship was citric and fixed acidity, with a correlation of almost 0.7. In terms of quality, the strongest relationship was seen with alcohol, and especially for the higher quality wines the trend was evident.


# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}
ggplot(aes(x=sulphates,y=alcohol,color= factor(quality)),data=rw) + 
  geom_point(position="jitter",
             alpha = 0.7,
             size = 3) + 
  scale_x_continuous(limits = c(0.3,1.4),
                     breaks = seq(0.3,1.4,0.1)) + 
  scale_y_continuous(limits= c(9,14),
                     breaks = seq(9,14,0.5)) + 
  scale_colour_brewer()
```

This plot shows us that in general, the high quality wines have both high sulphate and alcohol, which we knew from before. We also see that there's a tendency for the alcohol level to deviate relatively more than the sulphate when it comes to varying with quality.

```{r echo=FALSE,warnings=FALSE}

ggplot(aes(x=alcohol,color = factor(round_any(sulphates,0.2))),data=rw) + 
  geom_density(alpha = 0.25) +
  facet_wrap(factor('quality')) + 
  scale_color_brewer(palette="Reds")
```

This density plot shows us how the alcohol varies within different levels of quality and sulphates. We see how wines in both the ends of the quality scale have mostly low sulphate levels.  

```{r echo=FALSE,warnings=FALSE}
ggplot(aes(x=citric.acid,y=fixed.acidity,color= factor(round_any(volatile.acidity,0.2))),data=rw) + 
  geom_point(position="jitter",alpha = 0.7,size = 3) + 
  scale_x_continuous(limits = c(0,0.8),
                     breaks = seq(0,0.8,0.1)) + 
  scale_y_continuous(limits= c(6,16),
                     breaks = seq(6,16,1)) + 
  scale_colour_brewer(palette="Blues")
```

This plot shows the three acidity measurements.

Let's try a linear model with the most important chemical components to predict quality:
```{r echo=FALSE}
m1 <- lm(quality ~ alcohol, data = rw)
m2 <- update(m1, ~ . + volatile.acidity)
m3 <- update(m2, ~ . + sulphates)
m4 <- update(m3, ~ . + pH)
mtable(m1, m2, m3,m4)
rw["pred"] <- predict(m4,rw)


ggplot(aes(x=pred),data=rw) + geom_histogram(binwidth=0.1)+
  scale_x_discrete(breaks=seq(2,9,1)) + xlab ("Predicted quality")
```

This histogram shows what quality our linear model predicted for the original wines. We see that our model mostly places wines in the 4.5 to 7 range, which is not surprising given the distribution of the qualities. Even though our quality is given as a discrete and not continuous variable, the linear regression model gives us a hint of how a wine will score. Let's do a rounding, and have a look at how often we're missing and how much.

```{r echo=FALSE}
#Rounding the prediction to compare it with actual quality
rw["predRounded"] <- round(rw["pred"])

#Finding the difference between the rounded prediction and actual quality
rw["diff"]<- rw["quality"] - rw["predRounded"]

#Combine new dataframes to plot in same histogram
linearpreds <- data.frame(rw["predRounded"])
linearpreds$type<-"Prediction"
actual <- data.frame(rw["quality"])
actual$type<-"Wine tasters"

names(linearpreds) <- c("Quality","Type")
names(actual) <- c("Quality","Type")

combo <- rbind(linearpreds,actual)


ggplot(data=combo,
       aes(x=factor(Quality),fill=Type)) + 
  geom_histogram(position="dodge") +
  xlab("Quality") 
```
This histogram compares the predicted quality (when the predictions are rounded to the nearest integer), compared to the actual quality given by the wine tasters. 

```{r}

ggplot(aes(x=factor(diff)),data=rw) + 
  geom_histogram(binwidth=1,
                 fill="blue",
                 alpha=0.8) +
  scale_x_discrete(breaks=seq(-3,3,1)) +
  xlab("Difference between actual and predicted quality")
```

Here, I've taken the difference between the prediction and the actual quality. This doesn't look that bad, we have more zeroes than anything else, but remember that the distribution is already pretty even. Let's have a look at how the predictor quality = 5 performs.

```{r echo=FALSE}
rw["five"] <- 5
rw["diffive"] <- rw["quality"] - rw["five"]

lp <- data.frame(rw["diff"])
lp$type<-"Prediction"
five <- data.frame(rw["diffive"])
five$type<-"Default to 5"

names(lp) <- c("Difference","Type")
names(five) <- c("Difference","Type")

diffs <- rbind(lp,five)

ggplot(data=diffs,
       aes(x=factor(Difference),fill=Type)) + 
  geom_histogram(position="dodge") +
  xlab("Quality") 
```

The blue color indicates the count of wines that were assigned the correct quality by the rounded linear prediction. The red color indicates the count of wines that were assigned the correct quality by the naive prediction that all wines are quality 5. It's clear that the linear predictor has more correct predictions, but the difference here isn't huge.


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?
The strongest predictors for wine quality are alcohol and sulphates. However, there doesn't seem to any particular relation between the two. What we see is that high quality wines have a high alcohol level, while low quality wines have a low alcohol level. The sulphate level is in the lower mid range for at both the extremes of quality.

### Were there any interesting or surprising interactions between features?
Citric acid and fixed acidity had some relationship between them. Grouping by volatile acidity didn't show any easily recognizable pattern, but you could also tell they weren't purely random either. 

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.
I decided to have a look at a linear model based on alcohol, volatile acidity, sulphates and pH. I rounded the result to the nearest whole number, and it didn't perform horribly. It was however only slightly better than the trivial model median(quality), and so I wouldn't put too much emphasis on it.


# Final Plots and Summary

### Plot One
```{r echo=FALSE, warnings=FALSE,Plot_One}
ggplot(aes(x=citric.acid,
           y=fixed.acidity,
           color = factor(quality)),
           data=rw) + 
  geom_point() + 
  scale_colour_brewer(palette = "YlOrRd") + 
  stat_smooth(method="lm", 
              formula= y ~ poly(x,2),
              color = "black",
              size = 1,
              se = FALSE) + 
  scale_x_continuous(limits=c(0,0.75),
                     breaks=seq(0,0.75,0.1)) + 
  xlab("Citric acid (g /dm^3)") + 
  ylab("Fixed acidity (g/dm^3)") + 
  ggtitle("Fixed acidity vs. citric acid")
```

### Description One
The relationship between fixed acidity and citric acid seems to follow a quadratic polynomial, with a higher citric acid being related to an even higher fixed acidity. While fixed acidity varies between 6 and 16 $g/dm^3$, the ammount of citric acid varies between 0 and 0.8 $g/dm^3$. As to why the relationship shows the particular non-linear relation, one could speculate that the chemical process resulting in citric acid also produces a larger ammount of other acids.

### Plot Two
```{r echo=FALSE, Plot_Two,warnings=FALSE}
ggplot(aes(x=factor(quality),
           y=alcohol),
       data=rw) + 
  geom_boxplot(outlier.shape = NA,
               fill = "grey") +
  geom_point(position = "jitter",
             alpha = 0.25,
             color = "red") + 
  ylab("Alcohol (% by volume)") + 
  xlab("Quality") + 
  ggtitle("Alcohol versus quality")

```

### Description Two
All low quality wines have low alcohol levels. All high quality wines have high alcohol levels, increasing from quality five thru 8. The lowest average is found for quality 5 wines, at 9.9%. For quality 8 the average is 12.1. Perhaps the clearest sign of something goin on is the fact that 75% of all quality 8 wines have a higher alcohol percentage than all of the quality 3 wines. As mentioned earlier, this could be due to the age of the wines, which has not been given. Storing wine for a long time tends to increase alcohol. Another possibility is that good wines just tend to have a higher alcohol, or that wine tasters like the taste.

### Plot Three
```{r echo=FALSE, Plot_Three,warning=FALSE}
ggplot(aes(x=sulphates,color = factor(quality)),data=rw) + 
  geom_density(size=1.1,
               alpha = 0.7) + 
  scale_fill_brewer(palette = "Set1") + 
  scale_x_continuous(name = "Sulphates (g/dm^3)", 
                     limits = c(0.2,1.3),
                     breaks = seq(0.2,1.3,0.1)) + 
  scale_colour_brewer(palette="Blues") + 
  ggtitle("Sulphate per quality") + 
  ylab("Density")
```

### Description Three
This plot shows how the densities of the sulphate curves for each of the different qualities look. Higher quality wines have more sulphate, but also somewhat flatter curves. The mean difference is relatively big, with quality 7 wines at 0.74 $g/dm^3$ having 29% than quality 3 wnies at 0.57 $g/dm^3$. Sulphates typically go along with strong smells, and so perhaps this correlation is seen because wines rich on sulphate may have a stronger and more complex aroma.

------

# Reflection
The data set contains 1600 wines, and while this is a lot of wines, it's not that many at the same time. For example there are only around 10 wines with quality 3, and no wines with quality 0,1,2, 9 or 10. There's also quite a bit of relevant information we don't have,  we don't know the age of the wines for example. Aging wine will typically let water vaporize, and in turn thus increase a lot of chemical concentrations. Are the best wines in general older? We can't tell. I'm pretty sure age would be a valuable data addition, but also acknowledge that lookin only at chemical values was the point of the data set. I wasn't expecting to find very high correlations between quality and chemicals, because a lot of the experience of drinking a good red wine is attributed to the different specific aromas. However there's clearly some correlation, as seen with alcohol, pH, sulphates and so on.

My attempt at a linear model didn't yield very good results. This could mean a few things: The data measured doesn't explain the quality that well, there's too much noise, or that the linear model isn't very good here. Several of the features I chose have some linear behaviour in them, but this typically doesn't hold for both ends of the quality spectrum (look at alcohol, sulphates, volatile acidity). The linear model will sacrifice accuracy in the extreme qualities where the sample size is small, and when the outliers don't follow the pattern we get a bad model in that area. The result is a model that never suggests the wine should have a quality of 3,4 or 8. When you match this with the fact that quality 5 and 6 don't really show many distinct patterns, you get a lot pretty large error. I'm however fairly certain that other models could yield pretty good results, such as the SVM which is mentioned in the text.